### 2023-11-07 Notes/optest_update.tex
Added simulation 2.

Suppose that we want to estimate $\theta = E[g(X, Y_1, Y_2)]$ for a known
function $g$. The proposed estimator is

\begin{align}\label{eq:prop}
  &\hat \theta_{prop} =\\ \nonumber
  &n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] + n^{-1} \sum_{i = 1}^n
  \frac{\delta_{10}}{\pi_{10}} (E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i]) 
  + n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} (E[g_i \mid X_i, Y_{2i}] - E[g_i \mid X_i]) \\ \nonumber
  &+ n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} (g_i - E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i, Y_{2i}] + E[g_i \mid X_i]).
\end{align}

The goal is the show that over all functions $b_1(X, Y_1)$ and $b_2(X, Y_2)$,
$\hat \theta_{prop}$ is the optimal estimator in the form:

@@

@@

\[L_{2r}(G_r(Z)) = E[g(Z) \mid G_r(Z)].\]

@@

\subsection*{Linear Expectations}

@@

\subsection*{Projection onto Nuisance Tangent Space}

@@

\subsection*{Optimal Model}

Due to the fact that I do not know how to solve the previous equations, there 
is another way to construct an optimal model. We can find the values of 
$c_0$, $c_1$, and $c_2$ in $\hat \theta_c$ that minimize the variance the 
estimator. We can find these values by differentiating by $c_i$ and solving 
for $c_i$:

>>>>

Suppose that we want to estimate $\theta = E[g(X, Y_1, Y_2)]$ for a known
function $g$. 
The goal is the show that over all functions $b_1(X, Y_1)$ and $b_2(X, Y_2)$,
we have the optimal estimator in the form:

@@

%The proposed estimator is 
%
%\begin{align}\label{eq:propind}
%  &\hat \theta^{ind}_{prop} =\\ \nonumber
%  &n^{-1} \sum_{i = 1}^n E[g_i \mid X_i] + n^{-1} \sum_{i = 1}^n
%  \frac{\delta_{10}}{\pi_{10}} (E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i]) 
%  + n^{-1} \sum_{i = 1}^n \frac{\delta_{01}}{\pi_{01}} (E[g_i \mid X_i, Y_{2i}] - E[g_i \mid X_i]) \\ \nonumber
%  &+ n^{-1} \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} (g_i - E[g_i \mid X_i, Y_{1i}] - E[g_i \mid X_i, Y_{2i}] + E[g_i \mid X_i]).
%\end{align}

@@

\[L_{2r}(G_r(Z)) = c_r E[g(Z) \mid G_r(Z)]\]

where each $c_r$ is a constant. This actually suggests a variety of estimators 
that each have different values of coefficients of $\delta$. Initially, we tried 
a direct projection onto the nuisance tangent space, but I got stuck. See Appendix A 
for this work.

\subsection*{Different Classes of Estimators}

All of the considered estimators have a similar form:

\[\hat \theta = \frac{\delta_{11}}{\pi_{11}}g(Z) + \beta_0(\delta, c_0)E[g(Z)
\mid X] + \beta_1(\delta, c_1)E[g(Z) \mid X, Y_1] + \beta_2(\delta, c_2) E[g(Z)
\mid X, Y_2].\]

In this case, $c$ is a vector of constants $c = (c_0, c_1, c_2)$ that are 
chosen to minimize the variance of $\hat \theta$. See Appendix B for details on 
an example about choosing the values of $\hat c$.

\begin{tabular}{lrrr}
  \toprule
  Estimator & $\beta_0(\delta, c_0)$ & $\beta_1(\delta, c_1)$ & Implemented \\
  \midrule
  $\hat \theta_{prop}$ & $\left(1 - \frac{(\delta_{10} + \delta_{11})}{(\pi_{10} + \pi_{11})} - 
  \frac{(\delta_{01} + \delta_{11})}{(\pi_{01} + \pi_{11})} + \frac{\delta_{11}}{\pi_{11}}\right)$ &
  $\left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}} - \frac{\delta_{11}}{\pi_{11}}\right)$ & \checkmark \\
  $\hat \theta^{ind}_{prop}$ & $\left(1 - \frac{(\delta_{10})}{(\pi_{10})} - 
  \frac{(\delta_{01})}{(\pi_{01})} + \frac{\delta_{11}}{\pi_{11}}\right)$ &
  $\left(\frac{\delta_{10}}{\pi_{10}} - \frac{\delta_{11}}{\pi_{11}}\right)$ & \checkmark \\
  $\hat \theta_{c}$ & $c_0\left(1 - \frac{(\delta_{10} + \delta_{11})}{(\pi_{10} + \pi_{11})} - 
  \frac{(\delta_{01} + \delta_{11})}{(\pi_{01} + \pi_{11})} + \frac{\delta_{11}}{\pi_{11}}\right)$ &
  $c_1\left(\frac{\delta_{10} + \delta_{11}}{\pi_{10} + \pi_{11}} - \frac{\delta_{11}}{\pi_{11}}\right)$ & \\
  $\hat \theta_{c, ind}$ & $c_0\left(1 - \frac{(\delta_{10})}{(\pi_{10})} - 
  \frac{(\delta_{01})}{(\pi_{01})} + \frac{\delta_{11}}{\pi_{11}}\right)$ &
  $c_1\left(\frac{\delta_{10}}{\pi_{10}} - \frac{\delta_{11}}{\pi_{11}}\right)$ & \checkmark \\
  $\hat \theta_{\delta}$ & $c_0\left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{00}}{\pi_{00}}\right)$ &
  $c_1\left(\frac{\delta_{11}}{\pi_{11}} - \frac{\delta_{10}}{\pi_{10}}\right)$ & \checkmark \\
  \bottomrule
\end{tabular}

@@

\section*{Simulation Studies}

\subsection*{Simulation 1}

To see which of these estimators is the best, we run a simulation study. In this 
study, we have the following variables:

\begin{align*}
  \begin{bmatrix} x \\ e_1 \\ e_2 \end{bmatrix} 
  &\stackrel{ind}{\sim} N\left(\begin{bmatrix} 0 \\ 0 \\ 0 \end{bmatrix}, 
  \begin{bmatrix} 1 & 0 & 0 \\ 0 & 1 & \rho \\ 0 & \rho & 1 \end{bmatrix}\right) \\
  y_1 &= x + e_1 \\
  y_2 &= \theta + x + e_2 \\
\end{align*}

Furthermore, $\pi_{11} = 0.4$ and $\pi_{00} = \pi_{10} = \pi_{01} = 0.2$.
The goal of this simulation study is to find $\theta = E[Y_2]$. In other words,
$g(Z) = Y_2$. There are several algorithms for comparison which are defined 
as the following:

\begin{align*}
  Oracle &= n^{-1} \sum_{i = 1}^n g(Z_i)\\
  CC &= \frac{\sum_{i = 1}^n \delta_{11} g(Z_i)}{\sum_{i = 1}^n \delta_{11}} \\
  IPW &= \sum_{i = 1}^n \frac{\delta_{11}}{\pi_{11}} g(Z_i)\\
\end{align*}

\begin{table}[ht!]
\caption{The true value of $\theta$ is 5. $\rho = 0.5$. The test statistic and 
p-value are from a one-sample t-test to see if the estimator is biased.}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
Algorithm & Bias & SD & Tstat & Pval\\
\midrule
Oracle & 0.001 & 0.044 & 1.546 & 0.061\\
CC & 0.002 & 0.058 & 1.549 & 0.061\\
IPW & 0.006 & 0.210 & 1.504 & 0.066\\
$\hat \theta_{prop}$ & 0.002 & 0.051 & 1.911 & 0.028\\
$\hat \theta_{prop}^{ind}$ & 0.002 & 0.091 & 0.983 & 0.163\\
$\hat \theta_{c, ind}$ & 0.001 & 0.074 & 0.784 & 0.216\\
$\hat \theta_{\delta}$ & 0.002 & 0.051 & 1.964 & 0.025\\
\bottomrule
\end{tabular}
\end{table}

The main takeaway from this simulation study is that $\hat \theta_\delta$
and $\hat\theta_{prop}$ have the smallest standard deviation among the 
non-oracle estimators.

\subsection*{Simulation 2}
We now test a second simulation study. The setup of this simulation study 
is the same as the previous simulation. The only difference is that we 
are interested in estimating $\theta = E[Y_1^2 Y_2]$. In other words, now 
we have $g(Z) = Y_1^2 Y_2$.

\begin{table}[ht!]
  \caption{The true value of $E[g(Z)]$ is 10. $\rho = 0.5$. The test statistic and 
p-value are from a one-sample t-test to see if the estimator is biased.}
\centering
\begin{tabular}[t]{lrrrr}
\toprule
Algorithm & Bias & SD & Tstat & Pval\\
\midrule
Oracle & 0.007 & 0.529 & 0.741 & 0.229\\
CC & 0.023 & 0.824 & 1.518 & 0.065\\
IPW & 0.031 & 0.915 & 1.846 & 0.032\\
$\hat \theta_{prop}$ & 0.011 & 0.635 & 0.912 & 0.181\\
$\hat \theta_{c, ind}$ & 0.013 & 0.662 & 1.060 & 0.145\\
$\hat \theta_{\delta}$ & 0.011 & 0.636 & 0.983 & 0.163\\
\bottomrule
\end{tabular}
\end{table}

Like the first simulation, $\hat \theta_{prop}$ and $\hat \theta_{\delta}$ 
performed the best.

\section*{Next Steps}

\begin{itemize}
  \item[1.] It turns out that \cite{tsiatis2006semiparametric} contains 
    a general technique for solving non-monotone missing data. They show 
    that the optimal estimator is found via the following double projection:

    \[L_2 = \Pi\left(\Pi\left(\frac{\delta_{11}}{\pi_{11}}g(Z) \mid C,
    G_C(Z)\right) \mid Z\right)\]

    I can show that $\hat \theta_\delta$ is NOT such a double projection. 
    However, I think it is worth exploring what such an estimator (especially 
    if we consider a linear estimator) looks like.
  \item[2.] Similar to the previous point, \cite{tsiatis2006semiparametric} 
    discuss what the optimal estimator of a non-monotone problem looks like 
    and they determine that it solves the estimating equation

    \[\sum_{i = 1}^n \mathcal{L}(\mathcal{M}^{-1}(g(Z)))\]

    where $\mathcal{M}(g(Z)) = E[\mathcal{L}(g(Z)) \mid Z]$ and 
    $\mathcal{L}(g(Z)) = E[g(Z) \mid C, G_C(Z)]$. The difficult part is 
    understanding $\mathcal{M}^{-1}$, but there is conveniently the fact that

    \[\mathcal{M}^{-1}(g(Z)) = \lim_{n \to \infty} \phi_{n + 1}(Z) \text{ where } 
      \phi_{n + 1}(Z) = (I - \mathcal{M}) \phi_n(Z) + g(Z) \text{ and }
    \phi_0(Z) = g(Z).\]

    I have been investigating what this estimating equation yields if we 
    approximate $\mathcal{M}^{-1}$ with $\phi_1$. However, I have not 
    finished this yet.
  \item[3.] A different method to find the projection onto $\Lambda_2$ 
    could be to try to minimize the (KL) divergence between 
    $L_2$ and $\frac{\delta_{11}}{\pi_{11}}g(Z)$.
\end{itemize}

\appendix

\section{Linear Expectations}

@@

\subsection{Projection onto Nuisance Tangent Space}

@@

Unfortunately, I am now stuck because it is unclear to me how to use any actual 
data to solve this problem.

\section{Solving for $\hat c$}

We can find the values of 
$c_0$, $c_1$, and $c_2$ in $\hat \theta_c$ that minimize the variance the 
estimator. We can find these values by differentiating by $c_i$ and solving 
for $c_i$:
